### 2018-07-20 reviewers_response.md
Edits to address reviews

For now, this are just notes, we should make this a polished response.

## Reviewer 1 (Stephanie Hicks)

> 1. In this workflow, the authors start with a count table. However, the majority of researchers will start with raw reads (e.g. a FASTQ file). It would be great if the author discussed current best practices for the quantification step of scRNA-seq data. Alternatively, the authors could point to other references that have already been developed.

We do refer to Aaron's workflow at the beginning, but we should make it more clear. We should also mention `scPipe`.

> 2. I would like to see the authors take advantage of the rich functionality and data exploration tools for cell- and gene-specific quality control (QC) introduced in low-level analysis workflows such as the one from Lun et al. (2016). Also, in this workflow, the authors create multiple SummarizedExperiment objects (e.g. one with only the top 1000 highly variable genes (HVGs), one with all genes, etc). This doesn’t seem efficient, especially with large single cell data sets such as the 1.3 million cells from embryonic mouse brains. I think both of these concerns can now be addressed with efforts such as the recently developed SingleCellExperiment Bioconductor object (https://github.com/drisso/SingleCellExperiment). For example, the authors could add a “USE” column in the gene- or cell-specific meta table to represent whether or not a particular gene in a particular cell met the filtering criteria applied. The authors could store W in the reduceDim assay of the SingleCellExperiment object.

We now make full use of the SingleCellExperiment class. We also show how to store the variance in the object and how to filter based on those values.

> 3. In ZINB-WaVE, the authors specify the number of dimensions for the low-dimensional space (K) to be K=50. Could the authors add more details for the reader explaining why they picked K=50 and describe situations in which a user would want to specify a higher or lower K? In particular, it would be useful to discuss computational time in terms of number of genes and cells. Also, it would be useful to note that if you only wanted to use ZINB-WaVE to remove known covariates for normalization, you can use K=0.

Note that a more comprehensive discussion on the role of K on computational time is elsewhere (the zinbwave paper). The question of how to choose K is a very important one, but I'm not sure a workflow is the right place to discuss it in details. We should definitely give more guidance to the user on how to pick K. Perhaps we should compute a few values and compare the results in terms of AIC/BIC.

> Minor comments:
> 1. When selecting the top 1000 HVGs, why do the authors not take into account the overall mean-variance relationship and only select genes based on the variance?

This is mostly for illustration purposes, but we do find in our experience that naively selecting most variable genes gives a good set of informative genes, perhaps because it selects on/off genes. We have expanded this section to add a discussion about different strategies that account for the structure of the data can be employed.

> 2. It would be great if the authors referenced other tools available for similar analyses currently available. For example there are several available packages for normalization of scRNA-seq data, such as calculating global scaling factors can be done with scran (https://bioconductor.org/packages/release/bioc/html/scran.html) or gene and cell-specific scaling factors using SCnorm (https://github.com/rhondabacher/SCnorm). Alternatively, users might want to try using relative transcript counts using Census (https://bioconductor.org/packages/release/bioc/html/monocle.html).

Definitely. We should mention these. Perhaps even use these methods instead of zinbwave's normalized counts.

## Reviewer 2 (Andrew McDavid)

Major comments:

1. There is something wrong with the data download link in the F1000 version so that I am unable to download these files and actually reproduce the workflow. I experimented a bit to see if I could figure out how to download the data anyways, but will reserve further evaluation of this submission until this issue can be resolved by the authors.

```{r}
urls = c("https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE95601&format=file&file=GSE95601%5FoeHBCdiff% "https://raw.githubusercontent.com/rufletch/p63-HBC-diff/master/ref/oeHBCdiff_clusterLabels.)
```

Not sure what's wrong with the F1000 version. Perhaps the links are not rendered correctly? Let's look into this and make sure that the data can be downloaded.

2a. This workflow will likely be out-of-date when the underlying packages transition to use SingleCellExperiment. This is actually a positive thing because many of the more opaque lines of code (involving subsetting ERCC genes, etc) will be more streamlined.

Correct. Let's look into keeping the ERCC spike-ins in the object.

2b. It requires installation of the development branch of bioconductor, which impacts the usefulness of the workflow to the average user. I expect the authors will revise this tutorial when Bioconductor 3.6 is released and use of the devel branch is no longer necessary. Additionally `slingshot` is an requirement, but currently only exists on github and no SHA1 provided. I hope that `slingshot` will be added as a bioconductor package shortly. In the meantime, a tag must be added to the git repo for the release being used in this workflow and instructions provided for how to install this tag. Additionally, the authors may wish to note that installation instructions for the packages will be provided at the end of the workflow so that someone proceeding sequentially will not be tripped up.

Slingshot is in Bioconductor now. Good point about the installation instructions being at the end. I think we can provide a short section on installation at the beginning.

2c. Opaque code is presented in order to generate plots, e.g.
```{r}
palDF <- ceObj@clusterLegend[[1]]
pal <- palDF[, "color"]
names(pal) <- palDF[, "name"]
pal["-1"] = "transparent"
plot(fit$points, col = pal[primaryClusterNamed(ceObj)], main = "", pch = 20, xlab = "Component1", ylab = "Component2") legend(x = "topleft", legend = names(pal), cex = .5, fill = pal, title = "Sample")
```

While this complexity may be necessary, perhaps some of it could be encapsulated as accessor functions in the package? Too much complexity here may cause users to miss the forest for the trees.

The plotReduceDim() function now addresses this.

The authors could better motivate (or at least explain the impact of) some of the default parameters and procedures.
Why do we set a zcut threshold of 3 for the `scone` filtering? 
Why K=50 for zinbwave?
RSEC parameters

How should a user decide on a value for these parameters?

These are all great points. We should address them in more deatails.

## Reviewer 3 (Mike Love)

> It would be useful to put a link to the source code (or to the section where the link to the source code exists) near the top of the document.

We have modified the section "Package version" to include a link to the source code as well as instructions on how to install the packages needed by the workflow.

> I was confused a bit by "the first major bifurcation in the HBC lineage trajectory occurs prior to cell division". Can you be more specific about what you are referring to here by cell division, as without knowledge of the system, I'm not sure where the cell division you refer to should appear.

Thank you, we agree that the paragraph taken out of context is unclear. The first bifurcation identified by slingshot (occurring right after the clusters marked by DeltaHBC1 and DeltaHBC2 in Figure 2) produces two lineage trajectories: the first one (purple curve in Figure 2) generates sustentacular cells through a differentiation process characterized by the absence of cell cycle and division. To simplify, each HBC stem cells gives rise to one sustentacular cell. The other lineage trajectory (blue and orange curves in Figure 2) generates proliferative GBCs, which are characterized by cell cycle and division. These, in turn, generate olfactory sensory neurons and microvillous cells. We hope that this more detailed explanation will be useful to better understand our system. We have updated the text to add this more detailed explanation.

> "within a single object": It may be good to explain what an "object" here is. You could, for example, refer to Figure 2 of the Bioconductor Nature Methods paper.

We have added a paragraph discussing what a SingleCellExperiment object is and what are the advantages of using a single object throughout a data analysis workflow.

> Misspelling: "reasonnable"

Fixed, thanks.

> On filtering for most variable genes, I understand this decision, and I also recommend it during workshops before making ordination plots. I know that students are not always certain why we care about variance (unsupervised). I like to mention that these are the genes where the "action" is. A side point, the log(x+1) is not variance stabilizing for RNA-seq counts in general. This filter can give higher priority to low count genes than to genes where there is interesting biological variability (though I do not doubt that the very high biological variance genes will be preserved). It might be useful to show a vsn::meanSdPlot() for the matrix log1p(assay(se))?

Thank you, this is a very good point. See also comment by Reviewer 1 about modeling the mean-variance relationship. We have expanded this section to add a discussion about why we care about the most variable genes and how different strategies that account for the structure of the data can be employed.

> "correcting for batch effects": What are batch effects? (Of course, I know what they are, but a reader may not, and you could cite some of the single cell literature here.)

We have expanded our discussion on batch effects to include an explanation of what we mean by this term in this specific example and more generally.

> "Note that, in this case, the low-dimensional matrix W is not included in the computation of residuals to avoid the removal of the biological signal of interest.": I understood this sentence only on a second pass through. One problem is that you haven't defined W in the text yet (only in Figure 4). I would only reference the matrix W if you have defined it.

We agree, that the reference to the model is too specific. We have modified the sentence. [TODO -- still not sure how to best describe this step]

> Figure 6: Can you change the figure width so that PC1 is not squished?

Yes, Figure 6 has been changed.

> Is there a circularity to the recovery of published clusters in Figure 6? Was ZINB-WaVE used in Fletcher (2017)?

ZINB-WaVE was not used in the original paper, which used a more complex normalization strategy, based on regressing out a set of QC measures. One key result in favor of the ZINB-WaVE projection is that with this method slingshot is able to estimate the correct lineages with minimal supervision (only the starting cluster), while in the original paper, one of the terminal states had to be specified.

DR: Make sure this is true.

> Can you say what the meaning of the color white is in Figure 8 (in the text or caption near this figure)?

Yes, let's explain better the plot.

> Figure 15 refers back to Figure 2 but does not use the same color scheme for the known cell types, so the reader cannot verify if you've recovered the lineages from the publication. It would be good therefore to have a legend for these figures (Fig 15 and following) indicating which cell types the colors refer to (this information is in the unlabeled table above, but should be included as a legend here).

I think it would be easier to change the colors right after the table in which we compare to the published results and keep the color consistent for the rest of the workflow. Elizabeth, is there a wrapper function to change colors without manually changing the clusterLegend?

@Davide: Yes, there is a new function `recolorClusters` and `renameClusters` to change name or colors in a specific clustering. Also there is the `plotClusterLegend` function that could be used here to create a legend.

> Can you briefly describe what a GAM is ahead of Figure 17?

Yes, we can.